Example gene-wise analysis of the 1001 genomes data set

### 2018-01-05 example.Rmd
Added section about creating annotation tables.

```{r, echo = FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(warning = FALSE, collapse = TRUE, comment = "#>", tidy = TRUE, out.width = "90%", fig.height=4, fig.width=7, out.extra='style="margin: auto; display: block; padding-top: 15px;"')
library(r1001genomes)
library(DT)
library(dplyr)

@@

>>>>

```{r, echo = FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(warning = FALSE, collapse = TRUE, comment = "#>", tidy = TRUE, out.width = "90%", fig.height=4, fig.width=7, out.extra='style="margin: auto; display: block; padding-top: 15px;"')
library(r1001genomes)
library(DT)
library(dplyr)
library(reshape2)

@@

## Creating annotations

We can build a table of all of these structure function annotations in Excel as a CSV or within R. For the AFBs we will annotate the protein domains and a micro RNA binding site.

```{r}
# create matrix for F-box/LRR division in the coding sequence (CDS), 
#  based on Tan et al. supplementary alignment 
domains<-matrix(data = c(c(1,1,1,1,151,151,1), # Fbox start in CDS
                         3*c(51,47,46,46,96,96,58), # Fbox end in AA * 3 = CDS 
                         3*c(51,47,46,46,96,96,58)+1, # LRR start 
                         c(1785,1758,1728,1734,1872,1860,1779)), # LRR end 
             nrow = 4, ncol = 7, byrow = TRUE, 
             dimnames = list(c('Fbox_start', 'Fbox_end', 
                               'LRR_start', 'LRR_end'), 
                           c("TIR1","AFB1","AFB2", "AFB3",
                             "AFB4", "AFB5", "COI1")))

#create columns for miR393
miR393_start <- c(1519,1507,1504,1510,1657,1657,1551)
miR393_end <- miR393_start+18

# bind these columns to the domains
domains<-rbind(domains,miR393_start,miR393_end)

# convert to a data frame and convert values to AA numbering from CDS
domains <- as.data.frame(t(round(domains/3))) 
#domains
write.csv(domains, "manuscript/annotations_wide.csv")
```

You can used the `annotations_wide.csv` file as a basis to build annotations for your gene family. However because we can make alignments of our genes because they are so closely related it is not necessary to create an annotation for each gene in you family. Also this wide format doesn't make much sense if we are trying to annotate something unique to one gene. Instead I would recommend using a long format annotation. This will also make plotting these annotation on top of plots and alignments later much easier. Here is how we can convert the wide format annotation data frame into a long format.

```{r}
# create a column for the genes, which are currently the row names 
domains$gene <- row.names(domains)

# We'll order the gene names so that they will always be plotted in the 
# conventional order
domains$gene <- factor(x= domains$gene, levels = c('TIR1','AFB1','AFB2','AFB3','AFB4','AFB5','COI1'), ordered = T)

# Now we'll use melt() from the reshape2 package to "melt" the data frame into 
# long format. It's best to look at the output to figure out what melt does.
domains_long <- reshape2::melt(data = domains, 
                               id.vars = "gene", 
                               variable.name = "annotation", # column name for 
                               # all of the column names in the wide format
                               value.name = "position") # what is the name 
                               # of all of the values in the table
domains_long

# Let's add another column to indicate that these annotations are for 
# amino acid positions
domains_long$position_type <- "AA"
#write.csv(domains_long, "manuscript/annotation_long.csv")
```

I would use this `annotation_long.csv` to keep track of the structure/function information on individual genes you come across as you are reading the literature. Then you can map this information on top of the diversity plots and alignments that we will create below to see if these annotations apply to the whole family or just the single genes cited in the literature.
